# Reproducible Research: Peer Assessment 1

This data analysis is to process `activity.csv` dataset and will answer the following questions:

1. Whati is mean total number of steps taken per day?

2. What is the average daily activity pattern?

3. Fill-in the missing values in the dataset.

4. Are there differences in activity patterns between weekdays and weekends?

## Loading and preprocessing the data

The following R codes loaded the activity dataset and display the first several observations:

```{r}
activityData <- read.csv(file="activity.csv")
head(activityData)
```

## What is mean total number of steps taken per day?

First, we compute the total number of steps of each day:
```{r}
totalStepsEachDay <- tapply(activityData$steps, activityData$date, sum)
totalStepsEachDay
```

And we plot the histogram of the generated data and have an overall impression of the data. A `summary` command can provide the mean and median of the data.
```{r}
hist(totalStepsEachDay, breaks = 10)
summary(totalStepsEachDay)
```

## What is the average daily activity pattern?
The following command average out the daily steps at each interval:
```{r}
stepsPattern <- tapply(activityData$steps, activityData$interval, mean, na.rm = TRUE)
head(stepsPattern)
```

To extract the interval time from the dataset row names:
```{r}
interval <- as.integer(rownames(stepsPattern))
```

Thus, we can plot the average daily activity pattern:
```{r}
plot(interval, stepsPattern, type ="l")
```

## Imputing missing values
We can use `summary` command to identify how many `NA` values are in the dataset.
```{r}
summary(activityData)
```
There are 2304 `NA` in the dataset.  

We will fill the missing values with the average steps for that interval. And the new dataset will be stored in `activityDataNArm`.
```{r}
activityDataNArm <- activityData
indexNA <- which(is.na(activityDataNArm$steps))
activityDataNArm$steps[indexNA] <- stepsPattern[sapply(activityDataNArm$interval[indexNA], toString)]
```

Again, we compute the total number of steps of each day:
```{r}
totalStepsEachDayNArm <- tapply(activityDataNArm$steps, activityDataNArm$date, sum)
totalStepsEachDayNArm
```

And we plot the histogram of the newly-generated data. A `summary` command can provide the mean and median of the data.
```{r}
hist(totalStepsEachDayNArm, breaks = 10)
summary(totalStepsEachDayNArm)
```

By comparing the newly-generated data and the old data, we found that the median and mean didn't change. By using this method to fill the missing data, the frequency at the median increased.

## Are there differences in activity patterns between weekdays and weekends?
First, we should get the weekday labels for each observations and subsetting the data into two datasets: `weekdayData` and `weekendData`:

```{r}
# Get the weekday labels for each observation.
weekdayLabels <- weekdays(as.Date(activityDataNArm$date,format = "%Y-%m-%d"),abbreviate = TRUE)

# Get the index for the weekdays.
weekdayIndex <- which(weekdayLabels %in% c("Mon","Tue","Wed","Thu","Fri"))

# Get the index for the weekends.
weekendIndex <- which(weekdayLabels %in% c("Sat","Sun"))

# Subsetting the data
weekdayData <- activityDataNArm[weekdayIndex, ]
head(weekdayData)

weekendData <- activityDataNArm[weekendIndex, ]
head(weekendData)
```

And we can use exactly same method to compute the average activity patterns for both weekdays and weekend.
```{r}
# Get the weekday pattern.
patternWD <- tapply(weekdayData$steps, weekdayData$interval, mean, na.rm = TRUE)
head(patternWD)

# Get the weekend pattern.
patternWE <- tapply(weekendData$steps, weekendData$interval, mean, na.rm = TRUE)
head(patternWE)

# Plot the data
library(ggplot2)
library(gridExtra)
plot1 <- qplot(interval, patternWE, ylim = c(0,250), main = "Weekend", geom = "line")
plot2 <- qplot(interval, patternWD, ylim = c(0,250), main = "Weekday", geom = "line")
grid.arrange(plot1,plot2,nrow=2)
```


It looks like there is strong activity at early time of a day during weekday and the activities are more averaged during the weekends.